Cost of Living in Toronto

### 2024-09-27 paper/paper.qmd
last change - fixed typos

This report uses the "About Cost of living in Toronto for low-income households" [@citecityoftoronto] dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal. This data contains 17 variables on income and expenses such as `employment income`, `GST/HST credits`, `pension benefits`, `rent`, `food`, `transportation costs` for 11 different scenarios of a low-income household. The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc. A detailed explanation of this 11 household scenarios can be found in the Appendix. The dataset is also within a monthly time period, reflecting monthly income and expenses.

@@

# take the average
avg_income$"Average Income in $" <- rowMeans(subset(avg_income, select = c("Household_1",
                                                                           "Household_2",
                                                                           "Household_3",
                                                                           "Household_4",
                                                                           "Household_5",
                                                                           "Household_6",
                                                                           "Household_7",
                                                                           "Household_8",
                                                                           "Household_9",
                                                                           "Household_10",
                                                                           "Household_11"
                                                                           )))
avg_income <- avg_income %>% select("Item", "Average Income in $")

# rename the columns to something more presentable
colnames(avg_income) <- c("Type of Income", "Average Income in $")

#round to the nearest integer
avg_income$"Average Income in $" <- round(avg_income$"Average Income in $", 0)
avg_income %>% head(13) %>% knitr::kable()
```

@@

avg_expense %>% head() %>% knitr::kable()
```
## Difference between Income and Expense

@@

>>>>

This report uses the "About Cost of living in Toronto for low-income households" [@citeopendatatoronto] dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal. This data contains 17 variables on income and expenses such as `employment income`, `GST/HST credits`, `pension benefits`, `rent`, `food`, `transportation costs` for 11 different scenarios of a low-income household. The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc. A detailed explanation of this 11 household scenarios can be found in the @sec-appendix . The dataset is also within a monthly time period, reflecting monthly income and expenses.

@@

# take the average
avg_income$"Average Income ($)" <- rowMeans(subset(avg_income, select = c("Household_1",
                                                                           "Household_2",
                                                                           "Household_3",
                                                                           "Household_4",
                                                                           "Household_5",
                                                                           "Household_6",
                                                                           "Household_7",
                                                                           "Household_8",
                                                                           "Household_9",
                                                                           "Household_10",
                                                                           "Household_11"
                                                                           )))
avg_income <- avg_income %>% select("Item", "Average Income ($)")

# rename the columns to something more presentable
colnames(avg_income) <- c("Type of Income", "Average Income ($)")

#round to the nearest integer
avg_income$"Average Income ($)" <- round(avg_income$"Average Income ($)", 0)
avg_income %>% head(13) %>% knitr::kable()
```

@@

avg_expense %>% head(7) %>% knitr::kable()
```
## Difference between Income and Expense

@@

\newpage
